---
title: "Plotting IRI"
output: github_document
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(magrittr)
library(ggplot2)

cambodia_structure <- suppressMessages(
  # read data
  readr::read_csv("data-raw/iri_cambodia.csv")) %>%
  # make column names R friendly
  janitor::clean_names() %>%
  dplyr::mutate(type = dplyr::if_else(type == "N/A", NA_character_, type)) %>%
  dplyr::rename(percent_n = percent_no, percent_freq = percent_frq) %>%
  # make percentages percentages (between 0 and 1)
  dplyr::mutate_at(dplyr::vars(dplyr::contains("percent")), ~ ./100)

cambodia_iri <- cambodia_structure %>%
  # make sure "types" are ordered by iri
  dplyr::mutate(type = forcats::fct_reorder(type, iri, .desc = T)) %>%
  # remove types with zero freq
  dplyr::filter(percent_freq > 0)

# Calculate some variables for the plot
cambodia_iri_plot_df <- cambodia_iri %>%
  dplyr::mutate(ymax = percent_weight, 
                ymin = -percent_n, 
                xmax = cumsum(percent_freq), 
                xmin = dplyr::lag(xmax, default = 0), 
                xbreaks = (xmin + xmax)/2)

# Number of types to show
n_sp <- 6
cambodia_iri_plot_df %>%
  ggplot() +
  geom_rect(aes(xmin = xmin, xmax = xmax, ymin = ymin, ymax = ymax, fill = iri), 
            colour = "black") +
  # a dashed horizontal line on zero
  geom_hline(yintercept = 0, linetype = 2) +
  # Set x axis
  scale_x_continuous(breaks = cambodia_iri_plot_df$xbreaks[1:n_sp], 
                     labels = scales::percent(cambodia_iri_plot_df$percent_freq[1:n_sp]), 
                     minor_breaks = cambodia_iri_plot_df$xmax, 
                     sec.axis = sec_axis(trans = ~., 
                                         breaks = cambodia_iri_plot_df$xbreaks[1:n_sp], 
                                         labels = cambodia_iri_plot_df$type[1:n_sp])) +
  # Set y axis
  scale_y_continuous(labels = scales::label_percent()) +
  theme_minimal() +
  theme(axis.text.x.top = element_text(angle = 90), 
        legend.position = "none", 
        panel.grid.major.x = element_line(linetype = 2)) +
  
  labs(title = "IRI plot of a Cambodian fish community",
       x = "% frequency", y = "← % number ← | → % weight →")

